The finite representation property for some reducts of relation algebras

### 2020-12-21 filtration-stuff/Text.tex
K5 actually doesn't admit strict filtrations

\begin{theorem}
  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be an Euclidean model, $\Gamma$ a set of $\operatorname{Sub}$-closed formulas, and $\sim$ an equivalence relation having a finite index that respects $\Gamma$, then $\widehat{R} = (R^{min}_{\Gamma})^{E} \subseteq R^{max}_{\Gamma}$.

  Thus, ${\bf K}5$ admits strict filtrations.
\end{theorem}

\begin{proof}
  We put $R^{min}_{\Gamma} := \overline{R}$ for brevity.

  Recall that $\overline{R}^{E}$ has the form $\bigcup \limits_{n < \omega} (\overline{R})_n$, where
  \begin{enumerate}
    \item $\overline{R}_0 = \widehat{R}$
    \item $\overline{R}_{m + 1} = \overline{R}_n \cup ((\overline{R})^{-1} \circ \overline{R}_n)$
  \end{enumerate}
  One needs to show that for each $n < \omega$ $(\overline{R})_n \subseteq R^{max}_{\Gamma}$. We prove this by induction.
  Suppose $\mathcal{M}, y \models \varphi$ for $\Diamond \varphi \in \Phi$ and $\hat{x} \overline{R}^{E} \hat{x}$. We need $\mathcal{M}, x \models \Diamond \varphi$.
  \begin{enumerate}
    \item The case of $n = 0$ is trivial and it follows directly from the definition of a minimal filtration.
    \item Suppose $n = 1$. $\hat{x} \overline{R}_1 \hat{y}$ means that $(\hat{x}, \hat{y}) \in \overline{R} \cup (\overline{R}^{-1} \circ \overline{R})$.
    Let us assume that $(\hat{x}, \hat{y}) \in \overline{R}^{-1} \circ \overline{R}$. That is, there exists $\hat{z} \in W / \sim_{\Gamma}$ such that $(\hat{z}, \hat{x}), (\hat{z}, \hat{y}) \in \overline{R}$.

    $\hat{z} \overline{R} \hat{y}$ means that there are $z' \in \hat{z}$ and $y' \in \hat{y}$ such that $z' R y'$. $\mathcal{M}, y \models \varphi$ implies $\mathcal{M}, y' \models \varphi$.
    Then $\mathcal{M}, z' \models \Diamond \varphi$. $\hat{z} \overline{R} \hat{x}$ implies that there are $z'' \in \hat{z}$ and $x' \in \hat{x}$ with $z'' R x'$.

    $z'' \sim_{\Gamma} z'$ implies $\mathcal{M}, z'' \models \Diamond \varphi$. $\mathcal{M} \models {\bf K}5$, so $\mathcal{M}, z'' \models \Box \Diamond \varphi$. Thus, $\mathcal{M}, x' \models \Diamond \varphi$ since $x' \in R(z'')$. But $x' \in \hat{x}$, then $\mathcal{M}, x \models \Diamond \varphi$.

    \item Suppose $(\overline{R})_n \subseteq R^{max}_{\Gamma}$. Let us show that for $(\overline{R})_{n + 1}$. Then either $(\hat{x}, \hat{y}) \in (\overline{R})_n$ or $(\hat{z}, \hat{y}), (\hat{z}, \hat{x}) \in (\overline{R})_n$. The first alternative holds due to the IH. We consider only the second one.

    By Lemma..., we may visualise $(\hat{z}, \hat{y}), (\hat{z}, \hat{x}) \in (\overline{R})_n$ with the following sequences for some $\widehat{u_1}, \dots, \widehat{u_n}, \widehat{t_1}, \dots, \widehat{t_n}$:

    \xymatrix{
    &&& \hat{z} & \ar[l]_{\overline{R}} \widehat{u_1} \ar@{-}[r]^{R'} & \widehat{u_2} \ar@{-}[r]^{R'} & \dots \ar@{-}[r]^{R'} & \widehat{u_{n - 1}} \ar@{-}[r]^{R'} & \widehat{u_n} \ar[r]^{\overline{R}} & \hat{y} \\
    &&& \hat{z} & \ar[l]^{\overline{R}} \widehat{t_1} \ar@{-}[r]_{R'} & \widehat{t_2} \ar@{-}[r]_{R'} & \dots \ar@{-}[r]_{R'} & \widehat{t_{n - 1}} \ar@{-}[r]_{R'} & \widehat{t_n} \ar[r]_{\overline{R}} & \hat{x}
    }
    where $R'$ is either $\overline{R}$ or $\overline{R}^{-1}$

    $\mathcal{M}, y \models \varphi$ and $\widehat{u_n} \overline{R} \hat{y}$, so for some $u_n^{'} \in \widehat{u_n}$ and $y' \in \hat{y}$ with $u_n^{'} R y'$, so $\mathcal{M}, u_n^{'} \models \Diamond \varphi$ and $\widehat{\mathcal{M}}, \widehat{u_n} \models \Diamond \varphi$.

    We have either $\widehat{u_{n - 1}} \overline{R} \widehat{u_n}$ or $\widehat{u_n} \overline{R} \widehat{u_{n - 1}}$.

    If the first holds, then there are $u_{n - 1}^{'} \in \widehat{u_{n - 1}}$ and $u_{n}^{''} \in \widehat{u_{n}}$ with $u_{n - 1}^{'} R u_{n}^{''}$.
    $\mathcal{M}, u_{n}^{''} \models \Diamond \varphi$. wtf?

    If the second holds, then there $u_{n - 1}^{'} \in \widehat{u_{n - 1}}$ and $u_{n}^{''} \in \widehat{u_{n}}$ with $u_{n}^{''} R u_{n - 1}^{'}$. $\mathcal{M}, u_{n}^{''} \models \Diamond \varphi$ implies $\mathcal{M}, u_{n}^{''} \models \Box \Diamond \varphi$, so $\mathcal{M}, u_{n - 1}^{'} \models \Diamond \varphi$.

  \end{enumerate}
\end{proof}

\section{Filtration for {\bf K4}}

>>>>

\begin{theorem}
  ${\bf K}45$ admits strict filtrations.
\end{theorem}

\begin{proof}
  Let $\mathcal{M} = \langle W, R, \vartheta \rangle$ be a transitive Euclidean model and $\overline{\mathcal{M}} = \langle \overline{W}, \overline{R}, \overline{\vartheta} \rangle$ its minimal filtration through $\Gamma$, where $\Gamma$ is finite and Sub-closed. Let us put $\widehat{R} = \overline{R}^{+} \cup \overline{R}^{E}$. Let us show that $\overline{R}^{+} \cup \overline{R}^{E} \subseteq \overline{R}^{max}$.

  That is, if $\mathcal{M}, y \models \varphi$ for $\Diamond \varphi \in \Gamma$ and $\hat{x} \widehat{R} \hat{y}$, then $\mathcal{M}, x \models \Diamond \varphi$.

  Let $\hat{x} \widehat{R} \hat{y}$. Let us consider the case when $(\hat{x}, \hat{y}) \in \overline{R}^{E}$

  \begin{enumerate}
    \item Suppose $(\hat{x}, \hat{y}) \in \overline{R}$, then $\mathcal{M}, x \models \Diamond \varphi$ holds trivially by the definition of the minimal filtration.
    \item Suppose the statement holds $\overline{R}_n$ and $(\hat{x}, \hat{y}) \in \overline{R}_{n + 1} = \overline{R}_n \cup (\overline{R}_n^{-1} \circ \overline{R}_n)$. We consider the case of $(\hat{x}, \hat{y}) \in (\overline{R}_n^{-1} \circ \overline{R}_n)$.

    Then there exists $\hat{z}$ such that $(\hat{z}, \hat{x}), (\hat{z}, \hat{y}) \in \overline{R}_n$.

    By IH, $\mathcal{M}, z \models \Diamond \varphi$.

    $(\hat{z}, \hat{y}) \in \overline{R}_n$ iff there are $\hat{u}_1, \dots, \hat{u}_n$ such that

    \vspace{\baselineskip}

    \xymatrix{
    &&& \hat{z} & \ar[l]^{\widehat{R}} \hat{u}_1 \ar[r]_{\widehat{R}'} & \hat{u}_2 \ar[r]_{\widehat{R}'} & \dots \ar[r]_{\widehat{R}'} & \hat{u}_{n - 1} \ar[r]_{\widehat{R}'} & \hat{u}_n \ar[r]_{\widehat{R}} & \hat{y}
    }
    where $\widehat{R}'$ is either $\widehat{R}$ or $\widehat{R}^{-1}$.
  \end{enumerate}
  As it is known, $\Diamond \Diamond \varphi \to \Box \Diamond \varphi \in {\bf K}45$.

  \vspace{\baselineskip}

  $\hat{u}_1 \widehat \hat{z}$, that is, $u_1' R z'$ for some $u_1' \in \hat{u}_1$ and $z' \in \hat{z}$.
  That is, $\mathcal{M}, u_1' \models \Diamond \Diamond \varphi$, so $\mathcal{M}, u_1' \models \Diamond \varphi$ and $\overline{M}, \hat{u}_1 \models \Diamond \varphi$.

  We have $\hat{u}_1 \widehat{R}' \hat{u}_2$. Suppose $\mathcal{M}, u_1'' \models \Diamond \varphi$ and $u_1'' R u_2'$.
  We also have $\mathcal{M}, u_1'' \models \Box \Diamond \varphi$, thus, $\mathcal{M}, u_2' \models \Diamond \varphi$.

  Suppose $\hat{u}_2 \widehat{R} \hat{u}_1$ and $u_2' R u_1''$, then $\mathcal{M}, u_2' \models \Diamond \varphi$.

  Similarly, we have $\mathcal{M}, u_i \models \Diamond \varphi$ iff $\mathcal{M}, u_{i + 1} \models \Diamond \varphi$, whenever $\hat{u}_i \widehat{R}' \hat{u}_{i + 1}$.

  Finally, we have $\hat{u}_n \widehat{R} \hat{x}$. Thus, $u_n' R x'$ for some $u_n' \in \hat{u}_n$ and $x' \in \hat{x}$. $\mathcal{M}, u_n' \models \Diamond \varphi$, so $\mathcal{M}, u_n' \models \Box \Diamond \varphi$. Then $\mathcal{M}, x' \models \Diamond \varphi$.
\end{proof}

\begin{theorem}
  ${\bf K}5$ does not admit strict filtrations.
\end{theorem}

\begin{proof}
  Let us consider a ${\bf K}5$ model whose Euclidean closure of the minimal filtration does not give us a filtration.

  Let us consider a frame called $\mathcal{F}_{bad}$. We define this frame with the following graph:

\vspace{\baselineskip}

  \xymatrix{
  &&&&&&& w_5\ar@(ul,ur) \ar@<-.5ex>[d] \\
  &&&&&& w_1\ar@(ul,ur)  & w_4 \ar@<.5ex>[u] \\
  &&&&&& w_2 \ar[u] & w_3 \ar[u]
  }
\end{proof}

Let us define a valuation $\vartheta$ such that $\vartheta(p) = \{ w_5\}$ and $\vartheta(q) = \{ w_1\}$. Let us consider a minimal filtration of $\mathcal{M}_{bad}$ through the Sub-closure of $\Gamma = \{ \neg p, \neg \Diamond p\}$.

Clearly that $w_2 \sim_{\Gamma} w_3$, since $\neg p$ and $\neg \Diamond p$ are true both at $w_2$ and $w_3$.

Moreover, $R_{min} \cup (R_{min}^{-1} \circ R_{min})$ is not a subset of $R_{max}$ since $(\hat{w_1}, \hat{w_5}) \in (R_{min}^{-1} \circ R_{min})$, but $\Diamond p$ is not true at $w_5$.

Let us also note that strict filtrations of this model is not Euclidean.
Suppose by contrary that $\widehat{R}^{\mathcal{E}}$ is a strict filtraction of that model. So $R_{min}^E \subseteq \widehat{R}^{\mathcal{E}}$, since $R_{min}^E$ is the minimal Euclidean relation containing $R_{min}$. On the other hand, $R_{min}^E \not\subseteq R_{max}$, so is not $\widehat{R}^{\mathcal{E}}$.
\section{Filtration for {\bf K4}}

### 2020-12-21 frp-paper/Paper.tex
K5 actually doesn't admit strict filtrations

@@

\section{Axiomatisating ${\bf R}(+, ;)$.}

@@

Let $\mathcal{A}$ is a join-semilattice ordered semigroup, a subset $F \subseteq \mathcal{A}$ is called \emph{prime filter}, if $F$ is upwardly closed and $x + y \in F$ implies $x \in F$ or $y \in F$. The spectrum of $\mathcal{A}$, denoted as $\operatorname{Spec}(\mathcal{A})$, is the set of all prime filters.

\begin{definition} Let $\mathcal{A}$ be a join-semilattice ordered semigroup. A prenetwork over $\mathcal{A}$ is a pair $(E, V, N)$, where $E$ is a set of nodes, $V$ is a set of nodes (an equivalence relation on $E$), and $N$ is a function $l : V \to \operatorname{Spec}(\mathcal{A})$.

@@

  \begin{enumerate}
    \item Round $0$: $\forall$ picks a prime filter $F_0$ of $(\mathcal{A})$. $\exists$ responds with a prenetwork $\mathcal{N}_0 = (E_0 = \{ x_0, x_1 \}, V_0, l_0)$ such that the $l_0(x_0, x_1) = l_0(x_1, x_0) = F_0$ and $l_0(x_0, x_0) = l_0(x_1, x_1) = {\uparrow {\bf 1}}$.
    \item Round $n + 1$. Suppose, the prenetwork $\mathcal{N}_n = (E_n, V_n, l_n) \rangle$ has been played and $F_n \in \operatorname{Spec}(\mathcal{A})$.

    $\forall$ picks elements $a, b, c \in \mathcal{A}$ with $a \: ; \: b \geq c$

    $\exists$ responds with the network $\mathcal{N}_{n + 1} = (E_{n + 1}, V_{n + 1}, l_{n + 1})$
  \end{enumerate}

@@

  \begin{enumerate}
    \item If $\mathcal{A}$ is representable then $\exists$ have a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$.
    \item If $|\mathcal{A}| = \omega$ and $\exists$ have a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$ then $\mathcal{A}$ is representable.
    \item For each $n < \omega$ there exists a first-order sentence $\sigma_n$ such that $\exists$ have a winning strategy in $\mathcal{G}_n(\mathcal{A})$ iff $\mathcal{A} \models \sigma_n$.
  \end{enumerate}
\end{lemma}

@@

  \begin{enumerate}
    \item Let $h$ be a representation of $\mathcal{A}$
    \item
    \item
  \end{enumerate}
\end{proof}

@@

The author is sincerely grateful to Robin Hirsch, Ian Hodkinson, Stepan Kuznetsov, Valentin Shehtman, Eugeny Zolin, and his supervisor Ilya Shapirovsky for valuable conversations.\\The previous version contained an error. Thanks to Jas Semrl for noticing that.

>>>>

Let $\mathcal{A}$ is a join-semilattice ordered semigroup, a subset $F \subseteq \mathcal{A}$ is called \emph{prime filter}, if $F$ is upwardly closed, downwardly directed, and $x + y \in F$ implies $x \in F$ or $y \in F$. The spectrum of $\mathcal{A}$, denoted as $\operatorname{Spec}(\mathcal{A})$, is the set of all prime filters. Let $a \in \mathcal{A}$, then $\uparrow a$ is a prime filter generated by $a$.

\begin{lemma}
Let $\mathcal{A}$ be a join-semilattice ordered semigroup

  \begin{enumerate}
    \item Every element $a \in \mathcal{A}$ contains in some maximal filter,
    \item Every maximal filter is prime,
    \item If $a \not\leq b$ then there exist prime filter $F$ such that $a \in F$ and $b \not\in F$,
    \item If $a \in F_1$ and $b \in F_2$, where $F_1$ and $F_2$ are prime filters, then there exists a prime filter $F$ such that $F_1 \: \* \: F_2 \subseteq F$.
  \end{enumerate}
\end{lemma}

  \begin{enumerate}
    \item
    \item
    \item
    \item
  \end{enumerate}
\end{proof}

\subsection{Axiomatisating ${\bf R}(+, ;)$ and the finite representation property}

@@

@@

\begin{definition} Let $\mathcal{A}$ be a join-semilattice ordered semigroup. A prenetwork over $\mathcal{A}$ is a pair $(E, V, N)$, where $E$ is a set of nodes, $V$ is a set of nodes such that $\langle E, V \rangle$ is a complete directed graph, and $N$ is a function $l : V \to \operatorname{Spec}(\mathcal{A})$.

@@

  \begin{enumerate}
    \item Round $0$: $\forall$ picks elements $a, b$ with $a \not\leq b$. By Lemma, there exists a prime filter such that $a \not\in F_0$ and $b \in F_0$. $\exists$ responds with a prenetwork $\mathcal{N}_0 = (E_0 = \{ x_0, x_1 \}, V_0 = \{ (x_0, x_1)\}, l_0)$ such that $l_0(x_0, x_1) = F_0$.
    \item Round $n + 1$. Suppose, the prenetwork $\mathcal{N}_n = (E_n, V_n, l_n)$ has been played and $F_n \in \operatorname{Spec}(\mathcal{A})$.

    $\forall$ has the following two options:
    \begin{enumerate}
      \item ({\bf Composition move}): $\forall$ can pick $x, y, z \in E_n$ with $a \in l_n(x, y)$ and $b \in l_n(y, z)$. Then $\exists$ responds with $\mathcal{N}_{n + 1} = (E_{n + 1}, V_{n + 1}, l_{n + 1})$ such that $\mathcal{N}_{n + 1}$
      is the same as $\mathcal{N}_n$, but $l_{n + 1}(x, z) = F_{n+1}$, where $F_{n+1}$ is a prime filter that contains
      $l_{n}(x, z) \cup \uparrow(a \: ; \: b)$
      \item ({\bf Witness move}):

      $\forall$ picks nodes $x, y \in E_n$ and $a, b \in \mathcal{A}$ such that
    \end{enumerate}
  \end{enumerate}

@@

  \begin{enumerate}
    \item If $\mathcal{A}$ is representable then $\exists$ have a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$.
    \item If $|\mathcal{A}| = \omega$ and $\exists$ have a winning strategy in $\mathcal{G}_{\omega}(\mathcal{A})$ then $\mathcal{A}$ is representable.
  \end{enumerate}
\end{lemma}

@@

  \begin{enumerate}
    \item Let $h$ be a representation of $\mathcal{A}$
    \item
  \end{enumerate}
\end{proof}

\begin{definition}
  A term network
\end{definition}

\begin{lemma}
  For each $n < \omega$ there exists a first-order sentence $\sigma_n$ such that $\exists$ have a winning strategy in $\mathcal{G}_n(\mathcal{A})$ iff $\mathcal{A} \models \sigma_n$.
\end{lemma}

\begin{proof}
\end{proof}

@@

The author is sincerely grateful to Robin Hirsch, Ian Hodkinson, Stepan Kuznetsov, Valentin Shehtman, Eugeny Zolin, and his supervisor Ilya Shapirovsky for valuable conversations.\\The previous version of this paper contained an error. Many thanks to Jas Semrl for noticing that.
